---
title: "lme4::lmer Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lme4::lmer Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette is intended to show the wide variety of `lme4::lmer` models that can be handled by {equatiomatic}. The output uses the notation from [Gelman and Hill](http://www.stat.columbia.edu/~gelman/arm/). If you notice any errors in the notation, please [file an issue](https://github.com/datalorax/equatiomatic/issues). Similarly, if you try to use {equatiomatic} with an lmer model and end up with an error (either in notation or code) I would *really* appreciate if you could file an issue with a reproducible example.

This vignette displays many of the models that are covered by the package tests. It also illustrates many of the features of {equatiomatic}, including

* Automatic detection of group-level predictor variables (and the level at which they predict)
* Handling of interactions, both within a level and across levels
* Handling of any number of higher-level (or cross-classified) grouping factors (i.e., number of levels or random effects is infinite)
* Some support for variance-covariance structure outside of standard unstructured specifications

## Setup
```{r setup}
library(equatiomatic)
library(lme4)
```

## Unconditional models
A basic two-level unconditional model:
```{r unconditional1}
um_hsb <- lmer(math ~ 1 + (1|sch.id), data = hsb)
extract_eq(um_hsb)
```

And a model with multiple levels:

```{r unconditional2}
um_long3 <- lmer(score ~ 1 + (1|sid) + (1|school) + (1|district), 
                data = sim_longitudinal)
extract_eq(um_long3)
```

## Models with predictors at level 1
```{r l1-preds}
lev1_hsb <- lmer(math ~ female + ses + minority + (1|sch.id), hsb)
extract_eq(lev1_hsb)
```

Note that in the above, the mean structure at level 1 is broken out into a separate line. You can override this with the `mean_separate` argument.
```{r mean_separate}
extract_eq(lev1_hsb, mean_separate = FALSE)
```

Similarly, just like with standard `lm` models, you can specify wrapping, and how many terms per line

```{r wrapping}
extract_eq(lev1_hsb, wrap = TRUE, terms_per_line = 2)
```

And one more example with multiple levels

```{r l1-preds2}
lev1_long <- lmer(score ~ wave + (1|sid) + (1|school) + (1|district),
                  data = sim_longitudinal)
extract_eq(lev1_long)
```

## Models with unstructured variance-covariance specifications

This should generally work regardless of the complexity.

```{r unstructured-vcv1}
hsb1 <- lmer(math ~ female + ses + minority + (minority|sch.id),
               hsb)
extract_eq(hsb1)
```

Notice that it correctly parses which variable is randomly varying here. We can also make it more complicated.

```{r unstructured-vcv2}
hsb2 <- lmer(math ~ female + ses + minority + (female + ses|sch.id),
               hsb)
extract_eq(hsb2)
```

And even really complicated. Note the model below gives a warning.

```{r unstructured-vcv3}
hsb3 <- lmer(math ~ female * ses + minority + 
                 (ses * female + minority|sch.id),
             hsb)
extract_eq(hsb3)
```

# Group level predictors

In the `sim_longitudinal` data that comes with the package, the only level 1 predictor is `wave`. The `group` and `treatement` variables are at the student level (level 2) and `prop_low` is at the school level. Let's also add a district level variable (just the average score for each district).

```{r group-level-preds-prep}
# calculate district means
dist_mean <- tapply(sim_longitudinal$score,
                      sim_longitudinal$district,
                      mean)

# put them in a df to merge
dist_mean <- data.frame(district = names(dist_mean),
                        dist_mean = dist_mean)

# create a new df with dist_mean added
d <- merge(sim_longitudinal, dist_mean, by = "district")
```

Now we can fit a model that should have predictors at every level. We'll allow `wave` to vary randomly at each level too.

```{r group-preds-model1}
group_preds_m1 <- lmer(score ~ wave + group + treatment  + prop_low + dist_mean +
                         (wave|sid) + (wave|school) + (wave|district),
                       data = d)
extract_eq(group_preds_m1)
```

## Interactions
Interactions with multilevel models can be tricky because they can be within or across levels, and the notation changes depending on whether the random effect for the lower level (in the interaction term) is specified as varying randomly within the given level. Luckily, {equatiomatic} handles all of this for you.

First, let's look at a model with interactions only at level 1.

```{r l1-interaction}
l1_hsb_int <- lmer(math ~ minority*female + (1|sch.id),
                     data = hsb)
extract_eq(l1_hsb_int)
```

And now an interaction at only level 2

```{r l2-interaction}
l2_hsb_int <- lmer(math ~ meanses*sector + (1|sch.id),
                     data = hsb)
extract_eq(l2_hsb_int)
```

But more complicated are cross level interactions. Here's a quick example.

```{r cross-level1}
cl_long1 <- lmer(score ~ wave*treatment + (wave|sid) + (1|school) + 
                        (1|district),
                 data = sim_longitudinal)
extract_eq(cl_long1)
```

Note that the `treatement` variable is shown as a predictor of the level 1 intercept and the level 1 slope. But if the slope is not randomly varying at this level, then the notation has to change.

```{r cross-level2}
cl_long2 <- lmer(score ~ wave*treatment + (1|sid) + (1|school) + 
                        (1|district),
                 data = sim_longitudinal)
extract_eq(cl_long2)
```

This works even for really complicated models, including three-way interactions that contain a cross-level interaction. For example

```{r cross-level3}
cl_long3 <- lmer(score ~ wave*group*treatment + wave*prop_low*treatment +
                  (wave|sid) + (wave|school) +
                  (wave + treatment|district),
                sim_longitudinal)
extract_eq(cl_long3)
```

## Alternative variance-covariance specifications
Finally, there is some support for alternative variance-covariance specifications. For example, you may want to specify a model with only the variance terms estimated, and not the covariances.

```{r vars-only}
hsb_varsonly <- lmer(math ~ minority*female + (minority*female || sch.id),
                       data = hsb)
extract_eq(hsb_varsonly)
```

Or maybe you want to group by the same thing multiple times. In this exact model I don't think this makes any sense, but there are cases where it can. Note that, again, this model produces a warning.

```{r doublegroup}
hsb_doublegroup <- lmer(math ~ minority*female + 
                              (minority|sch.id) + (female|sch.id),
                         data = hsb)
extract_eq(hsb_doublegroup)
```

And finally, you can have a mix of different things and it should generally still work.

```{r mix}
long_mixed_ranef <- lmer(score ~ wave +
         (wave||sid) + (wave|school) + (1|school) + (wave||district),
         sim_longitudinal)
extract_eq(long_mixed_ranef)
```

With that said, this is the part of the code that I would consider the most "experimental" at this point, so please do reach out if you have issues or run into use cases that are not supported.

## Future developments
There are still a few things I'm planning to implement that I haven't quite gotten around to yet. These include

* Implementation of `use_coef`. This is relatively straightforward and should be implemented soon.
* Implementation of other features currently included in the `lm` extraction, including `intercept`, and `raw_tex` 
* Inevitable bug fixes

The range of models that you can fit with `lme4::lmer` is huge, but I hope this will handle a wide range of models. I also recognize that some users may want a different notation for the models. At this point, I'm not planning to implement other notations, but that could change if there's enough demand for it. I'm open to any/all suggestions for how to improve the package, and would *love* pull requests to help support the package, big or small.
